---
title: "ANALYSIS of SURGICAL ACTIVITY"
author: "Dr. Andrea PECCATI"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    toc_depth: 3
    thumbnails: false
    lightbox: true
    gallery: false
    collapsed: false
---

```{=html}
<style>
@media print {
  /* 1. Hide menus, navigation bars, and logos */
  #sidebar, .book-summary, .book-header, .book-logo {
    display: none !important;
  }

  /* 2. Unlock the main container to allow page breaks */
  html, body, .book, .book-body, .page-wrapper, .page-inner, .section {
    display: block !important;
    height: auto !important;
    overflow: visible !important;
    position: static !important;
    margin: 0 !important;
    padding: 0 !important;
  }

  /* 3. Remove the rightward offset created by the sidebar */
  .book-body {
    left: 0 !important;
  }

  /* 4. Prevent charts/images from being split across two pages */
  img, canvas, .ggplot2 {
    page-break-inside: avoid !important;
  }
}
</style>
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	fig.height = 6,
	fig.width = 8,
	message = FALSE,
	warning = FALSE,
	dpi = 100,
	out.height = "600px",
	out.width = "800px"
)
load("data/ver_1/export.RData")
library(kableExtra)
library(tidyverse)

# Create plot theme
  my_plot_theme <- theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 22, hjust = 0.5, margin = margin(t = 10, b = 20)),
      axis.title = element_text(face = "bold", size = 12),
      axis.text  = element_text(face = "bold", size = 12),
      axis.line  = element_line(color = "black", size = 0.5),
      axis.ticks = element_line(color = "black")
    )
```

<a href="" onclick="window.print()" class="print-button" name="top">

Print report

</a>

```{=html}
<style>
.print-button {
  font-weight: bold;
  text-align: center;
}
@media print {
  .print-button {
    display: none !important;
  }
}
</style>
```

\
The following objects were **exported in .RData format** form the `surgical_analysis.R` file to be imported in this report:

- the `check_missing` function

- the `surgeries_total` dataframe

- the `surgeries_stats_wide` dataframe

- the `procedures_stats_wide` dataframe

## 1 - DATA CHECK

Before data analysis, a quick **data check was performed to search for NA values** (amount and %).

```{r}
check_missing(surgeries_total) |>
  rename(
    "Variable" = variable,
    "NA values" = missing,
    "%" = pct,
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1:3, width = "120px")
```

## 2 - PROCEDURES per YEAR

This is the report of the **procedures performed between `r min(surgeries_stats_wide$year)` and `r max(surgeries_stats_wide$year)`**, [**grouped by year**]{.underline}.

### 2.1 - Table

```{r}
surgeries_stats_wide |> 
  select(year, procedures) |>
  mutate(cumulative = cumsum(procedures)) |>
  rename(
    "Year" = year,
    "Procedures" = procedures,
    "Running" = cumulative
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1:3, width = "120px")
```

### 2.2 - Plot

```{r}
ggplot(
  data = surgeries_stats_wide,
  aes(x = year)
) +
geom_bar(
  aes(y = procedures),
  stat = "identity",
  fill = "lightgrey",
  color = "black",
  alpha = 0.6
) +
geom_text(
  aes(
    y = procedures,
    label = procedures
  ),
  vjust = -0.5,
  fontface = "bold",
  size = 4
) +
scale_x_continuous(
  name ="",
  breaks = seq(
    min(surgeries_stats_wide$year),
    max(surgeries_stats_wide$year),
    by = 1
  )
) +
scale_y_continuous(
  name = "",
  expand = expansion(mult = c(0, 0.05))
) +
labs(
  title = paste0(
    "PROCEDURES per YEAR (",
    min(surgeries_stats_wide$year),
    "-",
    max(surgeries_stats_wide$year),
    ")"
  )
) +
my_plot_theme
```

## 3 - PROCEDURES as LEAD SURGEON per YEAR

This is the report of the **procedures performed as lead surgeon between `r min(surgeries_stats_wide$year)` and `r max(surgeries_stats_wide$year)`**, [**grouped by year**]{.underline}.

⚠️ Data show **A CLEAR AND CONSTANTLY INCREASING % OF PROCEDURES PERFORMES AS LEAD SURGEON, to indicate a [constantly increasing surgical autonomy]{.underline}**[.]{.underline}

### 3.1 - Table

```{r}
surgeries_stats_wide |> 
  select(year, lead, pct_lead) |>
  mutate(pct_lead = sprintf("%.2f%%", pct_lead)) |>
  mutate(cumulative = cumsum(lead)) |>
  rename(
    "Year" = year,
    "As lead surgeon" = lead,
    "%" = pct_lead,
    "Running" = cumulative
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1:4, width = "120px")
```

### 3.2 - Plot

```{r}
ggplot(
  data = surgeries_stats_wide,
  aes(x = year)
) +
geom_bar(
  aes(y = pct_lead),
  stat = "identity",
  fill = "lightgrey",
  color = "black",
  alpha = 0.6
) +
geom_text(
  aes(
    y = pct_lead,
    label = paste0(round(pct_lead, 0), "%")
  ),
  vjust = -0.5,
  fontface = "bold",
  size = 5
) +
scale_x_continuous(
  name ="",
  breaks = seq(
    min(surgeries_stats_wide$year),
    max(surgeries_stats_wide$year),
    by = 1
  )
) +
scale_y_continuous(
  name = "",
  expand = expansion(mult = c(0, 0.05))
) +
labs(
  title = paste0(
    "% as LEAD SURGEON per YEAR (",
    min(surgeries_stats_wide$year),
    "-",
    max(surgeries_stats_wide$year),
    ")")
) +
my_plot_theme
```

## 4 - TOP 15 PROCEDURES

This is the report of the **top 15 procedures, performed between `r min(surgeries_stats_wide$year)` and `r max(surgeries_stats_wide$year)`**.

Codes and descriptions of ICD-9 procedures were taken from the [CMS.gov website](https://www.cms.gov/medicare/coding-billing/icd-10-codes/icd-9-cm-diagnosis-procedure-codes-abbreviated-and-full-code-titles).

### 4.1 - Table

```{r}
procedures_stats_wide |>
  select(procedure, procedures, short_descr)|>
  rename(
    "ICD-9 Code" = procedure,
    "Procedures" = procedures,
    "ICD-9 Short description" = short_descr
  ) |>
  kbl(align="ccl")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1, width = "150px") |>
  column_spec(2, width = "150px") |>
  column_spec(3, width = "300px")
```

### 4.2 - Plot

```{r}
ggplot(
data = procedures_stats_wide,
aes(
  x = reorder(short_descr, procedures),
  y = procedures)
) +
geom_col(
  fill = "lightgrey",
  color = "black"
) +
geom_text(
  aes(label = procedure),
  y = 0,
  hjust = -0.1,
  fontface = "bold",
  size = 4
) +
geom_text(
  aes(label = procedures),
  hjust = -0.5, 
  fontface = "bold",
  size = 4
) +
scale_x_discrete(
  name = ""
) +
scale_y_continuous(
  name="",
  expand = expansion(mult = c(0, 0.05))
) +
labs(
  title = paste0(
    "TOP ",
    length(procedures_stats_wide$procedure),
    " ICD-9 PROCEDURES"
  ),
  caption = "Short description of ICD-9 procedures taken from CMS.gov"
) +
coord_flip() +
my_plot_theme
```
